---
title: "Foreign-born residents in the city of Madrid"
author: "miniesta4"
date: "February 5, 2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyr)
```

## Introduction

Time series on foreign population are periodically published by the Madrid City
Council in its data bank.

<http://www-2.munimadrid.es/CSE6/jsps/menuBancoDatos.jsp>

A small application have been built to show interactively the evolution of 
the number of people by geographic area and country of origin.

The application can be found at:

<https://miniesta4.shinyapps.io/madpoblacion/>

## Exploratory data analysis

The dataset is provided as a csv file. The variable "Fecha" is in columns. 
Some values of the variable "Pais" are not countries itself. There is also a row
for not foreign born residents.

```{r d}
fic <- "./MadPoblacion/Data/d.rds"
d <- readRDS(fic)
# str(d, give.attr = FALSE)
names(d)
d[1:5, 1:5]
```

## Data transformation

Some transformations have been made in order to obtain a tidy dataset.

```{r d1}
d1 <- pivot_longer(d, cols = -c(Pais, Area) , names_to = "Fecha",
                   values_to = "Num_habit")
d2 <- d1[complete.cases(d1), ]
d2$fecha2 <- as.Date(d2$Fecha, format = "%d/%m/%y")
d2[1:5, ]
```


## Source code repository

The source code for the application can be found following the next link:

[Repository](https://github.com/miniesta4/datasciencecoursera/tree/master/MadPoblacion)





