---
title: "El pueblo"
author: "miniesta4"
date: "January 20, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(leaflet)
```

## Map of Navarredondilla

This is the map of a very small village called Navarredondilla, 
located in the Alberche Valley, province of Ávila, Spain.  
The icon is the village's coat of arms. Beautiful in my opinion.  
The link goes to the web site of a weather station located at the center of village.   
Today is quite chilly up there.   
It is not bad for such a small village (200 inhabitants) to have this two things.  
I hope you like it.

```{r nav, echo=FALSE}
navIcon <- makeIcon(
  iconUrl = "/home/miniesta4/Pictures/Navarredondilla.JPG", 
  iconWidth = 50, iconHeight = 50,
  iconAnchorX = 50, iconAnchorY = 50
)

navSite <- "<a href='http://www.navarredondilla.info'>El tiempo en Navarredondilla</a>"


df <- data.frame(lat = 40.455, lng = -4.82)
df %>% leaflet() %>%
  addTiles() %>%
  addMarkers(icon = navIcon, popup = navSite)
```

