---
title: "Foreign-born residents in the city of Madrid"
author: "miniesta4"
date: "February 5, 2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

Time series on foreign population are periodically published by the Madrid City
Council in its data bank.

<http://www-2.munimadrid.es/CSE6/jsps/menuBancoDatos.jsp>

An small application have been build to show interactively the evolution of 
the number of people by geographic area and country of origin.

The application can be found at:

<>

## Exploratory data analysis

The dataset is a csv file. The variable "Fecha" is in columns and some values
of the variable "Pais" are not countries. Geographic area has been added.

```{r d}
# fic <- "./MadPoblacion/Data/d.rds"
# d <- readRDS(fic)
# dim(d)
# d[1:5, 1:5]
```

## Data transformation

Some transformations and removals of rows have been made in order to obtain a
tidy data set.

```{r d1}
# d1 <- pivot_longer(d, cols = -c(Pais, Area) , names_to = "Fecha", 
#                    values_to = "Num_habit")
# d2 <- d1[complete.cases(d1), ]
# d2$fecha2 <- as.Date(d2$Fecha, format = "%d/%m/%y")
# d2[1:5, ]
```


## Source code repository

The source code for the application can be found in the following link:

[Repository](https://github.com/miniesta4/datasciencecoursera/tree/develop/MadPoblacion)





